Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-05-28 cvx-flats-neurips21/cvx-flats.tex
Jessie tweaking pass over Sec 1-3

% examples of problems where we care a lot about prediction dimension (that happen to be finite discrete)
Despite the ubiquity of surrogate losses, we lack general frameworks to design and analyze consistent surrogates.
This state of affairs is especially dire when one seeks low \emph{prediction dimension}, the dimension of the surrogate prediction domain.
For example, in multiclass classification with $n$ labels, the prediction domain might be $\reals^n$.
In many type (I) settings, such as structured prediction and extreme classification, the prediction dimension of any convex and consistent surrogate often becomes intractably large, forcing one to sacrifice consistency for computational efficiency.
To understand whether this sacrifice is necessary, recent work developed tools like the feasible subspace dimension to lower bound the prediction dimension of any consistent convex surrogate~\citep{ramaswamy2016convex}.
Challenges of type (II) include estimating risk measures such as conditional value at risk (CVaR), with applications in financial regulation, robust engineering design, and algorithmic fairness.
Risk measures are not elicitable, meaning they cannot be specified via a target loss, and thus we seek a surrogate loss of low (or at least finite) prediction dimension.
Recent work~\citep{fissler2016higher,frongillo2015elicitation,frongillo2020elicitation} gives prediction dimension bounds for some of these risk measures, but without the requirement that the surrogate be convex; bounds for convex surrogates are left as a major open question.

We present a new tool, \emph{$d$-flats}, which unifies existing techniques to bound the convex consistency dimension in both settings above.
Using this tool, we give the first prediction dimension bounds for risk measures with respect to convex surrogates, addressing the open question for type (II) above.
In settings of type (I), $d$-flats recover and tighten the feasible subspace dimension result of \citet{ramaswamy2016convex}.
%In settings of type (I), $d$-flats give tighter bounds in some cases, e.g. feasible subspace dimension result of \citet{ramaswamy2016convex}.
% Applied to settings of type (1), we recover the feasible subspace dimension result of \citet{ramaswamy2016convex}, and give an example where our bound is even tighter.
Our framework rests on \emph{property elicitation}, a weaker and simpler condition than calibration, as a way to understand consistency across a wide variety of domains.
\raft{Old contributions commented out here}
% First, we formalize a notion of consistency with respect to a target statistic (Definition~\ref{def:consistent-prop}) and show its relationship to consistency with respect to a target loss (Lemma~\ref{lem:consistent-loss-implies-prop}).
% We then show indirect elicitation is a necessary condition for consistency (Theorem~\ref{thm:consistent-implies-indir-elic}).
% With these tools in hand, we present a new framework for deriving lower bounds on the prediction dimension of consistent convex surrogates (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) via indirect elicitation.
% These bounds are the first to our knowledge that can be applied in all four quadrants.
% Moreover, our framework can also give tighter bounds than previously existed in the literature.
% We illustrate this sharpness with new bounds for well-studied problems such as abstain loss (\S~\ref{sec:finite-calib}) and variance, CVaR, and other measures of risk and uncertainty (\S~\ref{sec:contin-consis}).

@@

    % edges
    \path[->,bend angle=10]
    (consistent) edge[<->] node[left]{\cite{bartlett2006convexity,steinwart2008support}} (calibrated)
    (consistent) edge[bend left] node[above, sloped]{Prop.~\ref{thm:consistent-implies-indir-elic}} (indir-elic)    
    (calibrated) edge[bend right] (indir-elic)
    (indir-elic) edge node[above, sloped]{Thm.~\ref{thm:convex-flats-inf-dim}} (d-flats)
    (d-flats) edge[bend right] node[below, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds)
    (d-flats) edge[bend left] node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
  \end{tikzpicture}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable condition (Theorem~\ref{thm:convex-flats-inf-dim}), based on $d$-flats, for the existence of a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results, while simultaneously applying to other settings (Q2,4).}
\label{fig:results-flow}
\end{figure}

@@

If given, we use $\ell: \R \times \Y \to \reals$ to denote a \emph{target loss}, with predictions $r\in\R$.
%where the goal is minimize expected loss, $\E ~ \ell(f(X),Y)$ for pairs $X,Y$ drawn from some distribution.
Similarly, $L: \reals^d \times \Y \to \reals$ will typically denote a \emph{surrogate} loss, with surrogate predictions $u \in \reals^d$.
In this case, $d$ is the prediction dimension of $d$.
We write $\L_d $ for the set of $\mathcal{B}(\reals^d) \otimes \Y$-measurable and lower semi-continuous surrogates $L : \reals^d \times\Y \to \reals$ such that $\E_{Y \sim p} L(u,Y) < \infty$ for all $u \in \reals^d, p \in \P$, that are minimizable in that $\argmin_{u} \exploss{L}{u}{p}$ is nonempty for all $p\in\P$ (see \S~\ref{app:minnable-assumption} for discussion).
Moreover, $\Lcvx_d \subseteq \L_d$ is the set of convex (in $\reals^d$ for every $y \in \Y$) losses in $\L_d$.
\btw{Sufficient condition for $\A$-normal convex integrand: $L$ is lower semi-continuous and $\mathcal{B}(\A) \otimes \Y$-measurable; $\E_p L(u,Y)$ finite for all $u, p$, there exists a $u_0$ for each $p$ so that $\E_p L(u_0, Y)$ is finite and continuous for Rockefellar's corollary, though it's stricter than Ioffe and Tokhimorov IIRC}
%Similarly, let $\L_d^{fin}$ be surrogates $L: \reals^d \times \Y \to \reals$ that are $\mathcal{B}(\reals^d) \otimes \Y$-measurable and minimizable, in the sense that $\argmin_{u} \exploss{L}{u}{p}$ is nonempty for all $p\in\P$-- we use this class of losses with finite $\Y$ in mind.
Set $\L = \cup_{d \in \mathbb{N}} \L_d$, and $\Lcvx = \cup_{d \in \mathbb{N}} \Lcvx_d$.
A loss $\ell: \R\times \Y \to \reals$ is \emph{discrete} if $\R$ is a finite set.
%A surrogate $L:\reals^d \times \Y \to \reals$ is \emph{convex} if $L(\cdot,y)$ is convex in $u$ almost surely in $\Y$.
%Typically, we notate finite report sets $\R$.

@@

\subsection{Property elicitation}\label{subsec:properties}
%\jessiet{This sentence might make me wary of minnability assumptions as a reviewer}
%\raft{True, but we address it below.  We could forward reference that discussion.  I think it's okay though.}
Arising from the statistics and economics literature, property elicitation is similar to calibration, but only characterizes exact minimizers of a surrogate~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting,lambert2009eliciting,lambert2018elicitation,frongillo2015vector-valued,frongillo2014general}.
Specifically, given a statistic or \emph{property} $\Gamma$ of interest, which maps a distribution $p \in \P \subseteq \simplex$ to the set of desired or correct predictions, the minimizers of $L$ should precisely coincide with $\Gamma$.
For example, squared loss $L(r,y) = (r-y)^2$ elicits the mean $\Gamma(p) = \E_p Y$.
%% Bo: we already mentioned this about p above. I think it's confusing here.
%For intuition, to relate to consistency, one can think of $p = \Pr[Y|X=x]$ as a conditional distribution, though the definition is also applied to point prediction settings.

@@

To relate property elicitation to consistency, we need to allow for a link function, which gives rise to the notion of \emph{indirect} elicitation.
For single-valued properties, this definition reduces to the natural requirement $\gamma = \psi \circ \Gamma$.
\begin{definition}[Indirect Elicitation]\label{def:indirectly-elicits}
	A surrogate loss and link $(L, \psi)$ \emph{indirectly elicit} a property $\gamma:\P \toto \R$ if $L$ elicits a property $\Gamma: \P \toto \reals^d$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	We say $L$ \emph{indirectly elicits} $\gamma$ if such a link $\psi$ exists.
  \btw{interesting discussion of set-valued properties commented out; revive later!}
\end{definition}

@@

%\citet[Corollary 10]{agarwal2015consistent} provide a necessary condition for the direct convex elicitation of single-valued properties, yielding bounds via the dimensionality of level sets.
Consistency is often the desired notion of corresponding to a prediction task since it is a prerequisite for deriving ERM bounds, but is often studied through the equivalent or necessary conditions of calibration, indirect elicitation, or embedding; notions of prediction dimension have been studied for surrogates satisfying each of the criteria.
\citet{ramaswamy2016convex} introduce \emph{convex calibration dimension} and some lower bonuds for problems in Quadrant 1 to study the prediction dimension of convex calibrated losses; in Q1, this is precisely equal to convex consistency dimension of a target loss.

@@

\paragraph{Q4: Variance}
Consider the task of estimating the variance $\Var(p) = \E_p[Y^2] - \E_p[Y]^2$.
Interestingly, the variance is not elicitable as its level sets are not convex~\cite{osband1985providing,lambert2008eliciting}, meaning this task falls under Quadrant 4.
The fact that the variance is not elicitable does not yield a lower bound on elicitation complexity of 2, as it does not rule out the variance being a link of a real-valued convex-elicitable property; cf.~\citet[Remark 1]{frongillo2020elicitation}.
In \S~\ref{app:omitted-examples}, we show $\eliccvx(\Var)=2$, meaning the lowest dimension of a convex loss to estimate conditional variance is 2. %~\citep{osband1985providing,lambert2018elicitation}.
This lower bound will follow from Theorem~\ref{thm:bayes-risk-lower-bound} in \S~\ref{sec:contin-consis} using the fact that variance is the Bayes risk of squared loss. % $L(r,y) = (r-y)^2$, which elicits the mean $\Gamma(p) = \E_p Y$.
Interestingly, while perhaps intuitively obvious, even this simple result is novel.

@@

Theorem~\ref{thm:convex-flats-inf-dim} subsumes the  bounds given by~\citet{ramaswamy2016convex} by showing that, if there is a $d$-flat through $p$ fully contained in a level set $\gamma_r$ (so we can apply Theorem~\ref{thm:convex-flats-inf-dim}) then the subspace of feasible directions at the same $p \in \C := \gamma_r$ has co-dimension at most $d$.

@@

\section{Reconstructing \citet[Thm.\ 16]{ramaswamy2016convex}}

>>>>

% examples of problems where we care a lot about prediction dimension (that happen to be finite discrete)
Despite the ubiquity of surrogate losses, we lack general frameworks to design and analyze consistent surrogates.
This state of affairs is especially dire when one seeks low \emph{prediction dimension}, the dimension of the surrogate prediction domain.
For example, in multiclass classification with $n$ labels, the prediction domain might be $\reals^n$.
In many type (I) settings, such as structured prediction and extreme classification, the prediction dimension of any convex and consistent surrogate often becomes intractably large, forcing one to sacrifice consistency for computational efficiency.
To understand whether this sacrifice is necessary, recent work developed tools like the feasible subspace dimension to lower bound the prediction dimension of any consistent convex surrogate~\citep{ramaswamy2016convex}.
Challenges of type (II) include estimating risk measures such as conditional value at risk (CVaR), with applications in financial regulation, robust engineering design, and algorithmic fairness.
\jessiet{cites?}
Risk measures are not elicitable, meaning they cannot be specified via a target loss, and thus we seek a surrogate loss of low (or at least finite) prediction dimension.
Recent work~\citep{fissler2016higher,frongillo2015elicitation,frongillo2020elicitation} gives prediction dimension bounds for some of these risk measures, but without the requirement that the surrogate be convex; bounds for convex surrogates are left as a major open question.

We present a new tool, \emph{$d$-flats}, which unifies existing techniques to bound the convex consistency dimension in both settings above.
Using this tool, we give the first prediction dimension bounds for risk measures with respect to convex surrogates, addressing the open question for type (II) above. \jessiet{above = \citet{frongillo2015elicitation}?}
In settings of type (I), $d$-flats recover and tighten the feasible subspace dimension result of \citet{ramaswamy2016convex}.
%In settings of type (I), $d$-flats give tighter bounds in some cases, e.g. feasible subspace dimension result of \citet{ramaswamy2016convex}.
% Applied to settings of type (1), we recover the feasible subspace dimension result of \citet{ramaswamy2016convex}, and give an example where our bound is even tighter.
Our framework rests on \emph{property elicitation}, a weaker and simpler condition than calibration, as a way to understand consistency across a wide variety of domains.
\raft{Old contributions commented out here}
% First, we formalize a notion of consistency with respect to a target statistic (Definition~\ref{def:consistent-prop}) and show its relationship to consistency with respect to a target loss (Lemma~\ref{lem:consistent-loss-implies-prop}).
% We then show indirect elicitation is a necessary condition for consistency (Theorem~\ref{thm:consistent-implies-indir-elic}).
% With these tools in hand, we present a new framework for deriving lower bounds on the prediction dimension of consistent convex surrogates (Corollaries~\ref{cor:Pcodim-flat-single-val-prop} and~\ref{cor:Pcodim-flat-elic-relint-prop}) via indirect elicitation.
% These bounds are the first to our knowledge that can be applied in all four quadrants.
% Moreover, our framework can also give tighter bounds than previously existed in the literature.
% We illustrate this sharpness with new bounds for well-studied problems such as abstain loss (\S~\ref{sec:finite-calib}) and variance, CVaR, and other measures of risk and uncertainty (\S~\ref{sec:contin-consis}).

@@

    % edges
    \path[->,bend angle=10]
    (consistent) edge[<->] node[left]{\cite{bartlett2006convexity,steinwart2008support}} (calibrated)
    (consistent) edge[bend left] node[above, sloped]{Prop.~\ref{thm:consistent-implies-indir-elic}} (indir-elic)    
    (calibrated) edge[bend right] (indir-elic)
    (indir-elic) edge node[above, sloped]{Thm.~\ref{thm:convex-flats-inf-dim}} (d-flats)
    (d-flats) edge[bend right] node[below, sloped]{Cor.~\ref{cor:Pcodim-flat-elic-relint-prop}} (q4-bounds)
    (d-flats) edge[bend left] node[above, sloped]{Cor.~\ref{cor:Pcodim-flat-single-val-prop}} (q1-bounds);
  \end{tikzpicture}
\caption{Flow and implications of our results. Compared to calibration, we suggest indirect elicitation as a simpler but almost-as-powerful necessary condition for consistency. In particular, we obtain a testable condition (Theorem~\ref{thm:convex-flats-inf-dim}), based on $d$-flats, for the existence of a $d$-dimensional consistent convex surrogate. This condition recovers and strengthens existing calibration-based results, while simultaneously applying to other settings (Q2,4).\jessie{Is there a way to emphasize the $\conscvx$ bounds for Q3, Q4?}}
\label{fig:results-flow}
\end{figure}

@@

If given, we use $\ell: \R \times \Y \to \reals$ to denote a \emph{target loss}, with predictions $r\in\R$.
%where the goal is minimize expected loss, $\E ~ \ell(f(X),Y)$ for pairs $X,Y$ drawn from some distribution.
Similarly, $L: \reals^d \times \Y \to \reals$ will typically denote a \emph{surrogate} loss, with surrogate predictions $u \in \reals^d$.
In this case, $d$ is the prediction dimension of $d$.
We write $\L_d $ for the set of $\mathcal{B}(\reals^d) \otimes \Y$-measurable and lower semi-continuous surrogates $L : \reals^d \times\Y \to \reals$ such that $\E_{Y \sim p} L(u,Y) < \infty$ for all $u \in \reals^d, p \in \P$, that are minimizable in that $\argmin_{u} \exploss{L}{u}{p}$ is nonempty for all $p\in\P$ (see \S~\ref{app:minnable-assumption} for discussion).\jessiet{Change order so it's App A?}
Moreover, $\Lcvx_d \subseteq \L_d$ is the set of convex (in $\reals^d$ for every $y \in \Y$) losses in $\L_d$.
\btw{Sufficient condition for $\A$-normal convex integrand: $L$ is lower semi-continuous and $\mathcal{B}(\A) \otimes \Y$-measurable; $\E_p L(u,Y)$ finite for all $u, p$, there exists a $u_0$ for each $p$ so that $\E_p L(u_0, Y)$ is finite and continuous for Rockefellar's corollary, though it's stricter than Ioffe and Tokhimorov IIRC}
%Similarly, let $\L_d^{fin}$ be surrogates $L: \reals^d \times \Y \to \reals$ that are $\mathcal{B}(\reals^d) \otimes \Y$-measurable and minimizable, in the sense that $\argmin_{u} \exploss{L}{u}{p}$ is nonempty for all $p\in\P$-- we use this class of losses with finite $\Y$ in mind.
Set $\L = \cup_{d \in \mathbb{N}} \L_d$, and $\Lcvx = \cup_{d \in \mathbb{N}} \Lcvx_d$.
A loss $\ell: \R\times \Y \to \reals$ is \emph{discrete} if $\R$ is a finite set.
%A surrogate $L:\reals^d \times \Y \to \reals$ is \emph{convex} if $L(\cdot,y)$ is convex in $u$ almost surely in $\Y$.
%Typically, we notate finite report sets $\R$.

@@

\subsection{Property elicitation}\label{subsec:properties}
%\jessiet{This sentence might make me wary of minnability assumptions as a reviewer}
%\raft{True, but we address it below.  We could forward reference that discussion.  I think it's okay though.}
Arising from the statistics and economics literature, property elicitation is similar to calibration, but only characterizes exact minimizers of a surrogate~\citep{savage1971elicitation,osband1985information-eliciting,lambert2008eliciting,lambert2009eliciting,lambert2018elicitation,frongillo2015vector-valued,frongillo2014general}.
Specifically, given a statistic or \emph{property} $\Gamma$ of interest, which maps a distribution $p \in \P \subseteq \simplex$ to the set of desired or correct predictions, the minimizers of $L$ should precisely coincide with $\Gamma$.
For example, squared loss $L(r,y) = (r-y)^2$ elicits the mean $\Gamma(p) = \E_p Y$.
\jessiet{Cut this sentence?  Or rephrase to be technically correct.  We bring up the mean with set notation right after the equation.}
%% Bo: we already mentioned this about p above. I think it's confusing here.
%For intuition, to relate to consistency, one can think of $p = \Pr[Y|X=x]$ as a conditional distribution, though the definition is also applied to point prediction settings.

@@

We relate property elicitation to consistency, and to do so we need to allow for a link function, which gives rise to the notion of \emph{indirect} elicitation.
For single-valued properties, this definition reduces to the natural requirement $\gamma = \psi \circ \Gamma$.
\begin{definition}[Indirect Elicitation]\label{def:indirectly-elicits}
	A surrogate loss and link $(L, \psi)$ \emph{indirectly elicit} a property $\gamma:\P \toto \R$ if $L$ elicits a property $\Gamma: \P \toto \reals^d$ such that for all $u \in \reals^d$, we have $\Gamma_u \subseteq \gamma_{\psi(u)}$.
	We say $L$ \emph{indirectly elicits} $\gamma$ if such a link $\psi$ exists.
  \btw{interesting discussion of set-valued properties commented out; revive later!}
\end{definition}

@@

%\citet[Corollary 10]{agarwal2015consistent} provide a necessary condition for the direct convex elicitation of single-valued properties, yielding bounds via the dimensionality of level sets.
Consistency is often the desired notion of corresponding to a prediction task since it is a prerequisite for deriving excess risk bounds, but is often studied through the equivalent or necessary conditions of calibration, indirect elicitation, or embedding; notions of prediction dimension have been studied for surrogates satisfying each of the criteria.
\citet{ramaswamy2016convex} introduce \emph{convex calibration dimension} and some lower bonuds for problems in Quadrant 1 to study the prediction dimension of convex calibrated losses; in Q1, this is precisely equal to convex consistency dimension of a target loss.

@@

\paragraph{Q4: Variance}
Consider the task of estimating the variance $\Var(p) = \E_p[Y^2] - \E_p[Y]^2$.
Interestingly, the variance is not elicitable as its level sets are not convex~\cite{osband1985providing,lambert2008eliciting}, meaning this task falls under Quadrant 4.
The fact that the variance is not elicitable does not yield a lower bound on elicitation complexity of 2, as it does not rule out the variance being a link of a real-valued convex-elicitable property; cf.~\citet[Remark 1]{frongillo2020elicitation}.
In \S~\ref{app:variance}, we show $\eliccvx(\Var)=2$, meaning the lowest dimension of a convex loss to estimate conditional variance is 2. %~\citep{osband1985providing,lambert2018elicitation}.
This lower bound will follow from Theorem~\ref{thm:bayes-risk-lower-bound} in \S~\ref{sec:contin-consis} using the fact that variance is the Bayes risk of squared loss. % $L(r,y) = (r-y)^2$, which elicits the mean $\Gamma(p) = \E_p Y$.
Interestingly, while perhaps intuitively obvious, even this simple result is novel.

@@

Theorem~\ref{thm:convex-flats-inf-dim} subsumes the  bounds given by~\citet{ramaswamy2016convex} by showing that, if there is a $d$-flat through $p$ fully contained in a level set $\gamma_r$ (so we can apply Theorem~\ref{thm:convex-flats-inf-dim}) then the subspace of feasible directions at the same $p \in \C := \gamma_r$ has co-dimension at most $d$, discussed in detail in \S~\ref{subapp:fsd-bound-reconsstruction}.

@@

\subsection{Reconstructing \citet[Thm.\ 16]{ramaswamy2016convex}}\label{subapp:fsd-bound-reconsstruction}
